Fourmile Calibrations

### 2023-02-20 calibration_output.Rmd
update spatial figs

@@

## Cumulative Tmean difference

2011-2020 average of daily station-grid Tmean for DOY 365 (or 366)

@@

#get rid of a bunch of data
hru_grid_tmean <- hru_grids %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_grid = tmean)

hru_statn_tmean <- hru_statn %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_statn = tmean)

tmean_hrus <- left_join(hru_grid_tmean, hru_statn_tmean, by = c('time', 'ID')) %>%
  mutate(diff = tmean_statn - tmean_grid,
         year = year(time)) %>%
  group_by(year, ID) %>%
  mutate(total_diff = cumsum(diff)) %>% 
  filter(time == max(time)) %>%
  group_by(ID) %>%
  summarize(tmean_avg_cumdiff = mean(total_diff)) %>%
  rename(value = ID)

@@

tm_shape(hrus_map_tmean) +
  tm_polygons('tmean_avg_cumdiff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=4)

@@

## Cumulative solar radiation difference

2011-2020 average of daily station-grid solar rad for DOY 365 (or 366)

@@

#get rid of a bunch of data
hru_grid_solrad <- hru_grids %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_grid = solRad)

hru_statn_solrad <- hru_statn %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_statn = solRad)

solrad_hrus <- left_join(hru_grid_solrad, hru_statn_solrad, by = c('time', 'ID')) %>%
  mutate(diff = solrad_statn - solrad_grid,
         year = year(time)) %>%
  group_by(year, ID) %>%
  mutate(total_diff = cumsum(diff)) %>% 
  filter(time == max(time)) %>%
  group_by(ID) %>%
  summarize(solrad_avg_cumdiff = mean(total_diff)) %>%
  rename(value = ID)

@@

tm_shape(hrus_map_solrad) +
  tm_polygons('solrad_avg_cumdiff', midpoint = NA, palette = 'YlOrRd', id = 'value') +
tm_shape(locations_solrad) +
  tm_symbols(col='blue', size=4)

@@

## Cumulative RH difference

@@

#get rid of a bunch of data
hru_grid_rh <- hru_grids %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_grid = rhum)

hru_statn_rh <- hru_statn %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_statn = rhum)

rh_hrus <- left_join(hru_grid_rh, hru_statn_rh, by = c('time', 'ID')) %>%
  mutate(diff = rh_statn - rh_grid,
         year = year(time)) %>%
  group_by(year, ID) %>%
  mutate(total_diff = cumsum(diff)) %>% 
  filter(time == max(time)) %>%
  group_by(ID) %>%
  summarize(rh_avg_cumdiff = mean(total_diff)) %>%
  rename(value = ID)

@@

tm_shape(hrus_map_rh) +
  tm_polygons('rh_avg_cumdiff', midpoint = NA, id = 'value') +
tm_shape(locations_rh) +
  tm_symbols(col='blue', size=4)

@@

## Cumulative wind difference

```{r wind, echo = F}

#get rid of a bunch of data
hru_grid_wind <- hru_grids %>%
  dplyr::select(c(time, ID, wind)) %>%
  rename(wind_grid = wind)

hru_statn_wind <- hru_statn %>%
  dplyr::select(c(time, ID, wind)) %>%
  rename(wind_statn = wind)

wind_hrus <- left_join(hru_grid_wind, hru_statn_wind, by = c('time', 'ID')) %>%
  mutate(diff = wind_statn - wind_grid,
         year = year(time)) %>%
  group_by(year, ID) %>%
  mutate(total_diff = cumsum(diff)) %>% 
  filter(time == max(time)) %>%
  group_by(ID) %>%
  summarize(wind_avg_cumdiff = mean(total_diff)) %>%
  rename(value = ID)

hrus_map_wind <- left_join(hrus, wind_hrus, by = 'value')

locations_w <- locations %>% 
  dplyr::filter(name %in% c('Gordon_South', 'Gordon_North',
                        'Betasso', 'Niwot_Snotel'))

tm_shape(hrus_map_wind) +
  tm_polygons('wind_avg_cumdiff', midpoint = NA, id = 'value') +
tm_shape(locations_w) +
  tm_symbols(col='blue', size=4)

>>>>

gw_maxcumsum <- gw %>%
  group_by(source, year) %>%
  summarize(yr_max_RD1 = max(RD1cumsum),
            yr_max_RD2 = max(RD2cumsum),
            yr_max_RG1 = max(RG1cumsum),
            yr_max_RG2 = max(RG2cumsum)) %>%
  mutate(year = as.factor(year)) %>%
  pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) + 
  ggtitle('Max annual cumulative sums')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  facet_wrap(~flow_type, scales = 'free') +
  ggtitle('Max annual cumulative sums')

@@

## Tmean difference

2011-2020 average of daily station-grid Tmean

@@

#get rid of a bunch of data
hru_grid_tmean <- hru_grids %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_grid = tmean)  %>%
  group_by(ID) %>%
  summarize(tmean_all_grid = mean(tmean_grid))

hru_statn_tmean <- hru_statn %>%
  dplyr::select(c(time, ID, tmean)) %>%
  rename(tmean_statn = tmean) %>%
  group_by(ID) %>%
  summarize(tmean_all_statn = mean(tmean_statn))

tmean_hrus <- left_join(hru_grid_tmean, hru_statn_tmean, by = c('ID')) %>%
  mutate(diff = tmean_all_statn - tmean_all_grid) %>%
  rename(value = ID)

@@

tm_shape(hrus_map_tmean) +
  tm_polygons('diff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=4)

## Tmax difference

```{r tmax, echo = F}

tmax_grid <- read_csv('C:/Users/sears/Documents/Repos/fourmile/HRU_output/reg_tmaxAll_prism.csv') %>% 
  pivot_longer(cols = `3`:`19`,
               names_to = 'value', 
               values_to = 'tmax_grid') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmax_mean_grid = mean(tmax_grid))

tmax_statn <- read_csv('C:/Users/sears/Documents/Repos/fourmile/HRU_output/reg_tmaxAll.csv') %>% 
  pivot_longer(cols = `3`:`19`,
               names_to = 'value', 
               values_to = 'tmax_statn') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmax_mean_statn = mean(tmax_statn))

tmax_hrus <- left_join(tmax_grid, tmax_statn, by = c('value')) %>%
  mutate(diff = Tmax_mean_statn - Tmax_mean_grid,
         value = as.numeric(value))

hrus_map_tmax <- left_join(hrus, tmax_hrus, by = 'value')

tm_shape(hrus_map_tmax) +
  tm_polygons('diff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=4)

## Tmax monthly average

```{r tmax month, echo = F}

tmax_grid_mo <- read_csv('C:/Users/sears/Documents/Repos/fourmile/HRU_output/reg_tmaxAll_prism.csv') %>% 
  pivot_longer(cols = `3`:`19`,
               names_to = 'value', 
               values_to = 'tmax_grid') %>%
  mutate(time = dmy(time),
         month = month(time)) %>%
  filter(month %in% c(3:6)) %>%
  group_by(value, month) %>%
  summarize(avg_tmax_grid = mean(tmax_grid))

tmax_statn_mo <- read_csv('C:/Users/sears/Documents/Repos/fourmile/HRU_output/reg_tmaxAll.csv') %>% 
  pivot_longer(cols = `3`:`19`,
               names_to = 'value', 
               values_to = 'tmax_statn') %>%
  mutate(time = dmy(time),
         month = month(time)) %>%
  filter(month %in% c(3:6)) %>%
  group_by(value, month) %>%
  summarize(avg_tmax_statn = mean(tmax_statn))

tmax_hrus_mo <- left_join(tmax_grid_mo, tmax_statn_mo, by = c('value', 'month')) %>%
  mutate(mo_tmax_diff = avg_tmax_statn - avg_tmax_grid,
         value = as.numeric(value))

tmap_mode('plot')

#march
tmax_hrus_mar<- tmax_hrus_mo %>%
  filter(month == 3)

hrus_map_tmax_mar <- left_join(hrus, tmax_hrus_mar, by = 'value')

mar_tmax <- tm_shape(hrus_map_tmax_mar) +
  tm_polygons('mo_tmax_diff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=1) +
  tm_layout(title = 'March avg tmax')

#april
tmax_hrus_apr<- tmax_hrus_mo %>%
  filter(month == 4)

hrus_map_tmax_apr <- left_join(hrus, tmax_hrus_apr, by = 'value')

apr_tmax <- tm_shape(hrus_map_tmax_apr) +
  tm_polygons('mo_tmax_diff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=1) +
  tm_layout(title = 'April avg tmax')

#may
tmax_hrus_may<- tmax_hrus_mo %>%
  filter(month == 5)

hrus_map_tmax_may <- left_join(hrus, tmax_hrus_may, by = 'value')

may_tmax <- tm_shape(hrus_map_tmax_may) +
  tm_polygons('mo_tmax_diff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=1) +
  tm_layout(title = 'May avg tmax')

#june
tmax_hrus_jun<- tmax_hrus_mo %>%
  filter(month == 6)

hrus_map_tmax_jun <- left_join(hrus, tmax_hrus_jun, by = 'value')

jun_tmax <- tm_shape(hrus_map_tmax_jun) +
  tm_polygons('mo_tmax_diff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=1) +
  tm_layout(title = 'June avg tmax')

tmap_arrange(mar_tmax, apr_tmax, may_tmax, jun_tmax)

## Tmin difference

```{r tmin, echo = F}

tmin_grid <- read_csv('C:/Users/sears/Documents/Repos/fourmile/HRU_output/reg_tminAll_prism.csv') %>% 
  pivot_longer(cols = `3`:`19`,
               names_to = 'value', 
               values_to = 'tmin_grid') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmin_mean_grid = mean(tmin_grid))

tmin_statn <- read_csv('C:/Users/sears/Documents/Repos/fourmile/HRU_output/reg_tminAll.csv') %>% 
  pivot_longer(cols = `3`:`19`,
               names_to = 'value', 
               values_to = 'tmin_statn') %>%
  mutate(time = dmy(time)) %>%
  group_by(value) %>%
  summarize(Tmin_mean_statn = mean(tmin_statn))

tmin_hrus <- left_join(tmin_grid, tmin_statn, by = c('value')) %>%
  mutate(diff = Tmin_mean_statn - Tmin_mean_grid,
         value = as.numeric(value))

hrus_map_tmin <- left_join(hrus, tmin_hrus, by = 'value')

tm_shape(hrus_map_tmin) +
  tm_polygons('diff', id = 'value') +
tm_shape(locations_t) +
  tm_symbols(col='blue', size=4)

@@

Relative Precip difference

```{r relative p, echo = F}

reg_p <- read_csv('C:/Users/sears/Documents/Repos/fourmile/HRU_output/reg_precipAll.csv',
                      skip = 21) %>%
  dplyr::select(-starts_with('@')) %>%
  mutate(time = dmy(time)) %>%
  summarize(across(`3`:`19`, ~ mean(.))) %>%
  pivot_longer(cols = `3`:`19`,
               names_to = 'value', 
               values_to = 'precip_avg') %>%
  mutate(value = as.numeric(value))

hrus_map_precip <- left_join(hrus_map_precip, reg_p, by = 'value')

hrus_map_precip <- hrus_map_precip %>%
  mutate(relative_avgcum_diff = precip_avg_cumdiff / precip_avg)

tm_shape(hrus_map_precip) +
  tm_polygons('relative_avgcum_diff', midpoint = NA, palette = 'YlGnBu', id = 'value') +
tm_shape(locations_p) +
  tm_symbols(col='blue', size=4)

## Solar radiation difference

2011-2020 average of daily station-grid solar rad

@@

#get rid of a bunch of data
hru_grid_solrad <- hru_grids %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_grid = solRad)  %>%
  group_by(ID) %>%
  summarize(solrad_all_grid = mean(solrad_grid))

hru_statn_solrad <- hru_statn %>%
  dplyr::select(c(time, ID, solRad)) %>%
  rename(solrad_statn = solRad) %>%
  group_by(ID) %>%
  summarize(solrad_all_statn = mean(solrad_statn))

solrad_hrus <- left_join(hru_grid_solrad, hru_statn_solrad, by = c('ID')) %>%
  mutate(diff = solrad_all_statn - solrad_all_grid) %>%
  rename(value = ID)

@@

tm_shape(hrus_map_solrad) +
  tm_polygons('diff', midpoint = NA, palette = 'YlOrRd', id = 'value') +
tm_shape(locations_solrad) +
  tm_symbols(col='blue', size=4)

@@

## RH difference

@@

#get rid of a bunch of data
hru_grid_rh <- hru_grids %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_grid = rhum)  %>%
  group_by(ID) %>%
  summarize(rh_all_grid = mean(rh_grid))

hru_statn_rh <- hru_statn %>%
  dplyr::select(c(time, ID, rhum)) %>%
  rename(rh_statn = rhum) %>%
  group_by(ID) %>%
  summarize(rh_all_statn = mean(rh_statn))

rh_hrus <- left_join(hru_grid_rh, hru_statn_rh, by = c('ID')) %>%
  mutate(diff = rh_all_statn - rh_all_grid) %>%
  rename(value = ID)

@@

tm_shape(hrus_map_rh) +
  tm_polygons('diff', midpoint = NA, id = 'value') +
tm_shape(locations_rh) +
  tm_symbols(col='blue', size=4)

@@

## Average soil sat difference comparison

```{r soilsat, echo = F}

#get rid of a bunch of data
hru_grid_soilsat <- hru_grids %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_grid = soilSat)  %>%
  group_by(ID) %>%
  summarize(soilsat_all_grid = mean(soilsat_grid))

hru_statn_soilsat <- hru_statn %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_statn = soilSat) %>%
  group_by(ID) %>%
  summarize(soilsat_all_statn = mean(soilsat_statn))

soilsat_hrus <- left_join(hru_grid_soilsat, hru_statn_soilsat, by = c('ID')) %>%
  rename(value = ID)

hrus_map_soilsat <- left_join(hrus, soilsat_hrus, by = 'value')

tmap_mode('plot')

soilavg_grid <- tm_shape(hrus_map_soilsat) +
  tm_polygons('soilsat_all_grid', midpoint = NA, id = 'value') +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average soil sat')

soilavg_statn <- tm_shape(hrus_map_soilsat) +
  tm_polygons('soilsat_all_statn', midpoint = NA, id = 'value') +
  tm_legend(legend.outside = T)

tmap_arrange(soilavg_grid, soilavg_statn)

## Average max soil sat comparison

```{r max soilsat, echo = F}

#get rid of a bunch of data
hru_grid_soilsat_max <- hru_grids %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_grid = soilSat)  %>%
  mutate(year = year(time)) %>% 
  group_by(ID, year) %>%
  summarize(soilsat_yrmax_grid = max(soilsat_grid)) %>%
  summarize(avg_max_grid = mean(soilsat_yrmax_grid))

hru_statn_soilsat_max <- hru_statn %>%
  dplyr::select(c(time, ID, soilSat)) %>%
  rename(soilsat_statn = soilSat) %>%
  mutate(year = year(time)) %>% 
  group_by(ID, year) %>%
  summarize(soilsat_yrmax_statn = max(soilsat_statn)) %>%
  summarize(avg_max_statn = mean(soilsat_yrmax_statn))

soilsat_hrus_max <- left_join(hru_grid_soilsat_max, hru_statn_soilsat_max, by = c('ID')) %>%
  rename(value = ID)

hrus_map_soilsat_max <- left_join(hrus, soilsat_hrus_max, by = 'value')

tmap_mode('plot')

soilavg_grid_max <- tm_shape(hrus_map_soilsat_max) +
  tm_polygons('avg_max_grid', midpoint = NA, id = 'value') +
  tm_legend(legend.outside = T) +
  tm_layout(title = 'Average max soil sat',
            title.position = c('left', 'top'))

soilavg_statn_max <- tm_shape(hrus_map_soilsat_max) +
  tm_polygons('avg_max_statn', midpoint = NA, id = 'value') +
  tm_legend(legend.outside = T)

tmap_arrange(soilavg_grid_max, soilavg_statn_max)

## Average ET comparison

```{r et, echo = F}
#get rid of a bunch of data
hru_grid_ET <- hru_grids %>%
  dplyr::select(c(time, ID, actET)) %>%
  rename(ET_grid = actET)  %>%
  group_by(ID) %>%
  summarize(ET_all_grid = mean(ET_grid))

hru_statn_ET <- hru_statn %>%
  dplyr::select(c(time, ID, actET)) %>%
  rename(ET_statn = actET) %>%
  group_by(ID) %>%
  summarize(ET_all_statn = mean(ET_statn))

ET_hrus <- left_join(hru_grid_ET, hru_statn_ET, by = c('ID')) %>%
  rename(value = ID)

hrus_map_ET <- left_join(hrus, ET_hrus, by = 'value')

tmap_mode('plot')

ETavg_grid <- tm_shape(hrus_map_ET) +
  tm_polygons('ET_all_grid', midpoint = NA, id = 'value') +
  tm_legend(legend.outside = T) +
  tm_layout(main.title = 'Average ET')

ETavg_statn <- tm_shape(hrus_map_ET) +
  tm_polygons('ET_all_statn', midpoint = NA, id = 'value') +
  tm_legend(legend.outside = T)

tmap_arrange(ETavg_grid, ETavg_statn)
